Democracy, Public Support, and Measurement Uncertainty

### 2021-01-03 paper/dcpo_demsupport.Rmd
add more work on intro

One important reason for these mixed results is the difficulty in measuring democratic support over time and across many countries.
Public support for democracy cannot be directly observed, and its incorrect measurement will limit inferences about the relationships between public opinion and institutional development.
Further, the survey data available across countries and over time on support for democracy--or indeed most topics in public opinion---are sparse and incomparable, greatly hindering broadly comparative research.
Recently, a few pioneering studies have sought to overcome the hurdle of sparse and incomparable data by developing latent variable measurement models of public opinion [see @Caughey2019; @Claassen2019; @Solt2020a].
A pair of prominent recent works took advantage of this latent variable approach to measure democratic support for over one hundred countries for up to nearly three decades and to then assess, respectively, its consequences for and roots in democratic change [@Claassen2020a; @Claassen2020b].
These works concluded, first and supporting the classic argument, that mass support had a positive impact on democratic change, especially the endurance of democracy [@Claassen2020a, 127-130], and, second and directly contrary to it, democratic change has a thermostatic effect on public support, that is, that rather than generating their own support, deepening democracies provoke a backlash [@Claassen2020b, 46-50].

The models employed in these studies' analyses, though, do not account for uncertainty in their measurement of democratic support.
Because they are unobserved, latent variables are inherently accompanied by a quantity of measurement uncertainty.
To leave this uncertainty unacknowledged is to make the implausible assumption that the latent variables are measured perfectly, an assumption which distorts both statistical and substantive inference [@Crabtree2015; @juhl_measurement_2019].

In this letter, we reexamine the classic arguments about support for democracy and democratic change tested in these two pieces while correcting this oversight.
In addition to incorporating the measurement uncertainty, we also sought to reduce it by employing a superior model of public opinion, the Dynamic Comparative Public Opinion (DCPO) model developed by @Solt2020a, and more data to estimate democratic support for 144 countries for up to 33 years between 1988 and 2020.
Our analyses reveal that the significant relationships between public support and democratic change disappear once measurement uncertainty is taken into account, both in replications of @Claassen2020a and @Claassen2020b with the studies' original data and in our extension DCPO analyses that incorporate additional data.
That is, once measurement uncertainty is accounted for, there is no empirical support for either claim: declining democratic support does not appear important to subsequent democratic backsliding, and changes in democracy do not spur a thermostatic response in democratic support.

These null results might be caused by theoretical, methodological reasons, or both.
Methodologically, the existing survey questions might not be able to capture mass democratic attitude due to a shortage of indicators [Solt, 2021??] and the ineffectiveness of indicators in measuring the multi-faceted democratic support [@wuttke2020have] and behaviorally democratic commitments [@graham_democracy_2020; @Carey2020; @McCoy2020].
Substantive explanations might also account for the insignificant relationship: the elite-driven democratic backsliding [@weingast_political_1997; @levitsky_how_2018], the lack of a consensus on democratic erosion between the public and experts [@carey_searching_2019], and the difference between the public support in abstract and public commitments in context [@bartels_ethnic_2020; @graham_democracy_2020; @McCoy2020; @prothro_fundamental_1960; @van2017myth
].

@@

@@

Lastly, verbal support in a pro-democratic stance is usually higher than behavioral commitments [@prothro_fundamental_1960]. 
Since there is a difference in public support for democracy between "in the abstract" and "under specific propositions" [@bartels_ethnic_2020], recent studies questioned whether conventional measures of public support for democracy could capture real public commitment to democracy. 
By using survey experiments, @graham_democracy_2020 pointed out a "fundamental blind spot" in traditional measures of public democratic support: general survey questions cannot capture voters' willingness and behavioral commitment to democratic principles in specific context where voters need to choose between democratic norms and their partisan identification or policy preference. 
It turned out that voters subordinated democratic norms to their partisan preference by not electorally penalizing their candidates who endorse measures that violate norms.
Furthermore, voters support/encourage democratic norm-eroding policies when their co-partisans are in power rather than only tolerating anti-democratic policies [@Carey2020].

>>>>

One important reason for these mixed results is the difficulty in measuring democratic support over time and across many countries.
Public support for democracy cannot be directly observed, and its incorrect measurement will limit inferences about the relationships between public opinion and institutional development.
Further, the survey data available across countries and over time on support for democracy--or indeed most topics in public opinion---are sparse and incomparable, greatly hindering broadly comparative research.
Recently, a few pioneering studies have sought to overcome the hurdle of sparse and incomparable data by developing latent variable measurement models of public opinion [see @Caughey2019; @Claassen2019; @Solt2020a].
A pair of prominent recent works took advantage of this latent variable approach to measure democratic support for over one hundred countries for up to nearly three decades and to then assess, respectively, its consequences for and roots in democratic change [@Claassen2020a; @Claassen2020b].
These works concluded, first and supporting the classic argument, that mass support had a positive impact on democratic change, especially the endurance of democracy [@Claassen2020a, 127-130], and, second and directly contrary to it, democratic change has a thermostatic effect on public support, that is that, rather than generating their own support, deepening democracies provoke a backlash and it is instead the erosion of democracy that calls forth greater public support [@Claassen2020b, 46-50].

The models employed in these studies' analyses, though, do not account for uncertainty in their measurement of democratic support.
Because they are unobserved, latent variables are inherently accompanied by measurement uncertainty.
To leave this uncertainty unacknowledged is to make the implausible assumption that the latent variables are measured perfectly, an assumption which distorts both statistical and substantive inference [see, e.g., @Crabtree2015; @juhl_measurement_2019].

In this letter, we reexamine the classic arguments about support for democracy and democratic change tested in these two pieces while correcting this oversight.
In addition to incorporating the measurement uncertainty, we also sought to reduce it by employing a superior model of public opinion, the Dynamic Comparative Public Opinion (DCPO) model developed by @Solt2020a, and more data to estimate democratic support for 144 countries for up to 33 years between 1988 and 2020.
Our analyses reveal that the significant relationships between public support and democratic change disappear once measurement uncertainty is taken into account, both in replications of @Claassen2020a and @Claassen2020b with the studies' original data and in our extension DCPO analyses that incorporate additional data.
That is, once measurement uncertainty is accounted for, there is no empirical support for either claim put forward in these two pieces: declining democratic support does not signal subsequent democratic backsliding, and changes in democracy do not spur a thermostatic response in democratic support.

There are many plausible explanations for these null results, both methodological and theoretical.
One possible methodological issue is that support for democracy is multidimensional, and the survey questions used to measure it---which ask respondents to assess the desirability or appropriateness of democracy, to compare democracy to some undemocratic alternative, or to assess one of these alternatives---simply do not capture all of the dimensions relevant to democratic change, such as commitments to political equality and freedom of expression [see @Schedler2007; @Wuttke2020].
Another is that the relationships among these survey items have varied substantially over time [see @wuttke2020have]; this would violate the assumptions of the public opinion models used and would yield faulty estimates of democratic support.
A third is that these survey questions do not ask respondents to prioritize democracy with regard to other values with which it may come into conflict, such as their partisanship, and so miss capturing the actual extent of support democracy would find among the public when it actually needed public support [see @Carey2020; @graham_democracy_2020; @McCoy2020].

From a theoretical perspective, the effect of democracy on public support may depend not on its mere existence but on its effectiveness [see, e.g., @Magalhaes2014] and particularly with regard to redistribution [see, e.g., @Krieckhaus2014].
Similarly, the effect of public support on democracy may depend on the extent to which those who support democracy are also dissatisfied with the current regime's performance; that is, that it is not democratic support that is important to democratic change, but rather unsatisfied democratic demand [see, e.g., @Qi2011].
Of course, it could in fact be that democratic change does not actually depend on public opinion of any sort at all, but rather is an elite-driven phenomenon [see, e.g., Przeworski1991, @levitsky_how_2018].

@@

At a time when democracy as seen as under threat around the world [e.g., @Diamond2015], taken together, Claassen (2020_a_,_b_) send what is ultimately a reassuring message: the fate of democracy rests with us, the public, and when democratic institutions are undermined, we will swing to their support and constitute "an obstacle to democratic backsliding" [@Claassen2020b, 51].
Both of these conclusions may well be true, but the evidence we have, properly assessed, does not provide support for them.
There can be no room for complacency.

@@

Lastly, verbal support in a pro-democratic stance is usually higher than behavioral commitments [@prothro_fundamental_1960]. 
Since there is a difference in public support for democracy between "in the abstract" and "under specific propositions" [@bartels_ethnic_2020], recent studies questioned whether conventional measures of public support for democracy could capture real public commitment to democracy. 
By using survey experiments, @graham_democracy_2020 pointed out a "fundamental blind spot" in traditional measures of public democratic support: general survey questions cannot capture voters' willingness and behavioral commitment to democratic principles in specific context where voters need to choose between democratic norms and their partisan identification or policy preference. "fail to capture voters’ willingness to act on their commitment to democracy precisely when democracy is at stake"
It turned out that voters subordinated democratic norms to their partisan preference by not electorally penalizing their candidates who endorse measures that violate norms.
Furthermore, voters support/encourage democratic norm-eroding policies when their co-partisans are in power rather than only tolerating anti-democratic policies [@Carey2020].
